A Framework for Discrete Change

### 2008-11-11 papers/individual/individual.tex
Updated the individual paper and removed unneccesary graphics files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@239 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\abstract{A class of models is developed for measuring and detecting
discrete change in learning and development.  The basic model for
detecting such change is the latent or hidden Markov model.
Traditionally, these models were restricted to categorical, mostly
binary, observed variables, placing severe restrictions on possible
measurement models.  In this paper, the basic model is extended to
include arbitrary distributions for the observed variables, including
multi-variate distributions.  Moreover, there is optional support to
include time-varying predictors.  In effect, this model consists of
mixtures of general linear models with Markovian depencies over time
to model the change process.  In addition, transition parameters can
be made to depend on covariates as well, such that the switching
regime between states depends on characteristics of the individual or
the experimental situation.  The model is illustrated with an example
of participants' learning in the weather prediction task.}

@@

% phenomena of discrete change in development/learning
Piagetian developmental theory (REFERENCE??)  assumes step-wise
changes in the strategies that children apply in all kinds of tasks
such as the conservation learning and the balance scale task.
\citet{Maas1992} developed a catastrophe model to describe phase
transitions in learning and developmental processes.  They applied the
catastrophe model to learning in the conservation of liquid task
(REFERENCE??)  in which children have to judge relative volumes of
liquid in glasses of different heights and widths.  Young children
tend to ignore the width dimension and hence always choose the glass
with the highest level of liquid (REFERENCE??).  \citet{Maas1992}
showed that there is a sudden transition to a new strategy in which
the children also take the width of the glasses into account when
judging the volume of liquids.

@@

% conditioning and addiction research
Also in animal learning and conditioning, evidence is found for sudden
changes in response behavior \citet{Gallistel2004}.  In particular,
in their study, evidence was found for sudden onset of learning: at
the start of the learning experiment, the pigeons did not learn
anything and performance was stable; after a number of trials,
learning kicks in and there are large increases in performance.  The
interest here is in modeling the distribution of onset times: that is,
the trials at which learning suddenly takes off.  A similar interest
in process onset times is found in addiction research.  For example,
(REFERENCE) study the age at which children start using alcohol and
how this related to eventual outcomes in terms of addiction.

@@

% latent Markov versus time series analyses/outline
In above mentioned applications, the data consist mostly of a few
repeated measurements administered to large groups of participants.
The focus in the current chapter is rather on data that consist of
many repeated measurements, or time series, observed in only a few
participants or even just a single participant.  For example,
\citet{Visser2009} analyzed data from a single participant in an
experimental task that manipulates the trade-off between speed and
accuracy.  The data consisted of three time series with each around
150 repeated measurements of both reaction time and accuracy.  Below
we provide examples of analyzing time series from single participants
from two experiments; one from the Iowa Gambling Task and one from the
weather prediction task.  The interest in these tasks is to show that
participants develop different strategies over time in responding to
the stimuli and that the transition from one strategy to the next is a
discrete event.  Before providing these illustrations, below we give a
formalization of dependent mixture models and a brief overview of the
DepmixS4 package that was developed to specify and fit such models.

@@

% outline
In this section we describe a class of models which are especially
suitable for describing and testing discrete change in (individual)
time series data.  The dependent mixture model is similar to, but
slightly different from two other types of models that are in use for
modelling discrete change: the hidden and the latent Markov models.

% history of Markov models/LMM applications
Markov models have been used extensively in the social sciences; for
example, in analyzing language learning \cite{Miller1952,Miller1963},
in the analysis of paired associate learning \cite{Wickens1982}.  In
these models, the focus is on survey type data: a few repeated
measurements taken in a large sample.  \citeA{Langeheine1990}
discusses latent Markov models and their use in sociology ands
political science (see also McCutcheon, 1987).  Latent transition
models, for example, been used in studying development of math skills
\citep{Collins1992} and in medical applications \citep{Reboussin1998};
\citet{Kaplan2008} provides an overview of such models, that are
called stage-sequential models in the developmental psychology
literature.

@@

% hidden Markov models
Hidden Markov models (HMM) tend to be used in the analysis of long
univariate and individual timeseries.  For example, HMMs are the model
of choice in speech recognition applications \cite{Rabiner1989}.  In
biology, HMMs are used to analyze DNA sequences and in econometric
science, to analyze changes in stock market prices and commodities
\cite{Kim1994}.

% depmix model
The dependent mixture model that we propose here spans the range from
latent Markov models for few repeated measurements with many
participants to hidden Markov models for individual time series.  In
addition, the dependent mixture model includes multivariate responses.
The dependent mixture model consists of the following elements:
\begin{enumerate}
	\item $S$ is a collection of discrete states
	\item $S_{t} = \mat{A}S_{t-1}+\xi_{t}, \mat{A}$, a transition matrix
	\item $O_{t} = \mat{B}(S_{t}) + \zeta_{t}, \mat{B}$,  an observation density
\end{enumerate}

% state space
The state space, which is a set of discrete states, captures the
different states that the learning or developmental process under
consideration can be in.  In the balance scale example mentioned
above, children are applying one of two possible strategies in
responding to the items.  The states are characterized by their
corresponding observation densities.  Using for example Rule 1 in the
balance scale task leads to correct answers on some items and
incorrect answers on others.  A different strategy may lead to correct
answers on some items and to guessing behavior on other items.  In
analyzing categorization learning data, in which participants learn to
categorize a set of objects, a typical initial state is that
participants are guessing because at the start of the task they have
no knowledge of which features are important in categorization.

@@

% log-likelihood
The log-likelihood of DMMs is usually computed by the
so-called forward-backward algorithm \citep{Baum1966,Rabiner1989}, or
rather by the forward part of this algorithm.  \cite{Lystig2002}
changed the forward algorithm in such a way as to allow computing the
gradients of the log-likelihood at the same time.  They start by
rewriting the likelihood as follows (for ease of exposition the
dependence on the model parameters is dropped here):
\begin{equation}
	L_{T} = Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) = \prod_{t=1}^{T}
Pr(\vc{O}_{t}|\vc{O}_{1},
	\ldots, \vc{O}_{t-1}),
	\label{condLike}
\end{equation}
where $Pr(\vc{O}_{1}|\vc{O}_{0}):=Pr(\vc{O}_{1})$. Note that for a
simple, i.e.\ observed, Markov chain these probabilities reduce to
$Pr(\vc{O}_{t}|\vc{O}_{1},\ldots,
\vc{O}_{t-1})=Pr(\vc{O}_{t}|\vc{O}_{t-1})$.
The log-likelihood can now be expressed as:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log[Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots,
\vc{O}_{t-1})].
	\label{eq:condLogl}
\end{equation}

To compute the log-likelihood, \cite{Lystig2002} define the following
(forward) recursion:
\begin{align}
	\phi_{1}(j) &:= Pr(\vc{O}_{1}, S_{1}=j) = \pi_{j} b_{j}(\vc{O}_{1})
	\label{eq:fwd1} \\
\begin{split}
	\phi_{t}(j) &:= Pr(\vc{O}_{t}, S_{t}=j|\vc{O}_{1}, \ldots,
\vc{O}_{t-1}) \\
	&= \sum_{i=1}^{N} [\phi_{t-1}(i)a_{ij}b_{j}(\vc{O}_{t})] \times
(\Phi_{t-1})^{-1},
	\label{eq:fwdt}
\end{split}
\end{align}
where $\Phi_{t}=\sum_{i=1}^{N} \phi_{t}(i)$. Combining
$\Phi_{t}=Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, \vc{O}_{t-1})$, and
equation~(\ref{eq:condLogl}) gives the following expression for the
log-likelihood:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log \Phi_{t}.
	\label{eq:logl}
\end{equation}

@@

% outline/abstract: this has redundant information, mostly discussed
% in detail below
\pkg{depmixS4} implements a general framework for defining and fitting
dependent mixture models in the R programming language \citep{R2008}.
This includes standard Markov models, latent/hidden Markov models, and
latent class and finite mixture distribution models.  The models can
be fitted on mixed multivariate data with multinomial and/or gaussian
distributions.  Parameters can be estimated subject to general linear
constraints.  Parameter estimation is done through an EM algorithm or
by a direct optimization approach with gradients using the
\code{Rdonlp2} optimization routine when contraints are imposed on the
parameters.

% design goals
The \pkg{depmixS4} package was motivated by the fact that Markovian
models are used commonly in the social sciences, but no comprehensive
package was available for fitting such models.  Common programs for
Markovian models include Panmark \citep{Pol1996}, and for latent class
models Latent Gold \citep{Vermunt2003}.  Those programs are lacking a
number of important features.  There are currently some packages in R
that handle hidden Markov models but they lack a number of features
that we needed in our research.  In particular, \pkg{depmixS4} was
designed to meet the following goals:
\begin{enumerate}

	\item to be able to fit transition models with covariates, i.e.,
	to have time-dependent transition matrices

	\item to be able to include covariates in the prior or initial
	state probabilities of models

	\item to allow for easy extensibility, in particular, to be able
	to add new response distributions, both univariate and
	multivariate, and similarly to be able to allow for the addition
	of other transition models, e.g., continuous time observation
	models

@@

Although \pkg{depmixS4} is designed to deal with longitudinal or time
series data, for say $T>100$, it can also handle the limit case with
$T=1$ in analyzing cross-sectional data.  In those cases, there are no
time dependencies between observed data, and the model reduces to a
finite mixture model, or a latent class model.  Although there are
other specialized (R) packages to deal with mixture data, one specific
feature that we needed which is not available in other packages is the
possibility to include covariates on the prior probabilities of class
membership.

\subsection{Response distributions and parameters}

% this needs some more on the design, interface to glm and similar
% other models etc.
The package is built using S4 classes (object oriented classes in R)
to allow easy extensibility (Chambers, 1998).

Each row of the transition matrix and the initial state probabilities:
\begin{itemize}
	\item is modeled as a multinomial distribution
	\item uses the logistic link function to include covariates
	\item can have time-dependent covariates
\end{itemize}

% response distributions include the possibility to have direct
% effects of covariates or predictors as well as indirect effects
% through the use of covariates on the transition matrix and the
% initial state probabilities
Current options for the response models are models from the
generalized linear modeling framework:

\begin{itemize}
	\item normal distribution; continuous, gaussian data
	\item binomial logistic; binary data
	\item Poisson distribution; count data
	\item multinomial logistic; multiple choice data
\end{itemize}

All response models have the option of including covariates
Other link functions may be used; eg the probit for binary data.

% possibility of multivariate responses: local independence or true
% multivariate distributions, eg multivariate normal

@@

Parameters are estimated in \pkg{depmixS4} using the EM algorithm or
through the use of a general Newton-Raphson optimizer.  The EM
algorithm however has some drawbacks.  First, it can be slow to
converge towards the end of optimization (although it is usually
faster than direct optimization at the start, so possibly a
combination of EM and direct optimization is fastest).  Second,
applying constraints to parameters can be problematic; in particular,
EM can lead to wrong parameter estimates when applying constraints.
Hence, in \pkg{depmixS4}, EM is used by default in unconstrained
models, but otherwise, direct optimization is done using \pkg{Rdonlp2}
\cite{Tamura2007,Spellucci2002}, because it handles general linear
(in)equality constraints, and optionally also non-linear constraints.

@@

\begin{itemize}
	\item depmixS4 can be downloaded from: http://r-forge.r-project.org/depmix/
	\item It is feasible to fit hidden Markov models in moderate length time series
	\item Many applications in experimental psychology
	\item Future developments:
	\begin{enumerate}
		\item richer measurement models, eg factor models, AR models etc
		\item richer transition models, eg continuous time measurement occasions
		\item explicit state durations
		\item identifiability of models
		\item model selection
	\end{enumerate}
	\item standard errors of parameters
\end{itemize}

>>>>

\abstract{A class of models is developed for measuring and detecting
discrete change in learning and development.  The basic model for
detecting such change is the latent or hidden Markov model.
Traditionally, these models were restricted to categorical, mostly
binary, observed variables, placing severe restrictions on possible
measurement models.  In this paper, the basic model is extended to
include arbitrary distributions for the observed variables, including
multi-variate distributions.  Moreover, there is optional support to
include time-varying predictors.  In effect, this model consists of
mixtures of general linear models with Markovian depencies over time
to model the change process.  In addition, transition parameters can
be made to depend on covariates as well, such that the switching
regime between states depends on characteristics of the individual or
the experimental situation.  The model is illustrated with an example
of participants' learning in the Weather Prediction Task and in the 
Iowa Gambling Task.}

@@

% phenomena of discrete change in development/learning
Piagetian developmental theory assumes step-wise changes in the
strategies that children apply in all kinds of tasks such as the
conservation learning and the balance scale task (REFERENCE??).
\citet{Maas1992} developed a catastrophe model to describe step-wise
changes or phase transitions in learning and developmental processes.
They applied the catastrophe model to learning in the conservation of
liquid task (REFERENCE??)  in which children have to judge relative
volumes of liquid in glasses of different heights and widths.  Young
children tend to ignore the width dimension and hence always choose
the glass with the highest level of liquid (REFERENCE??).
\citet{Maas1992} showed that there is a sudden transition to a new
strategy in which the children also take the width of the glasses into
account when judging the volume of liquids.

@@

% conditioning and addiction research
Also in animal learning and conditioning, evidence is found for sudden
changes in response behavior \citet{Gallistel2004}.  In particular, in
their study, evidence was found for sudden onset of learning: at the
start of the learning experiment, the pigeons did not learn anything
and performance was stable; after a number of trials, learning kicks
in and there are large increases in performance.  The interest here is
in modeling the distribution of onset times: that is, the trials at
which learning suddenly takes off.  A similar interest in process
onset times is found in addiction research.  For example,
\cite{Sher2004} study the age at which children start using alcohol
and how this related to eventual outcomes in terms of addiction.

@@

% latent Markov versus time series analyses/outline
In above mentioned applications, the data consist mostly of a few
repeated measurements administered to large groups of participants.
The focus in the current chapter is rather on data that consist of
many repeated measurements, or time series, observed in only a few
participants or even just a single participant.  For example,
\citet{Visser2009} analyzed data from a single participant in an
experimental task that manipulates the trade-off between speed and
accuracy.  The data consisted of three time series with each around
150 repeated measurements of both reaction time and accuracy.  Below
we provide examples of analyzing time series from single participants
from two experiments; one from the Iowa Gambling Task and one from the
Weather Prediction Task.  The interest in these tasks is to show that
participants develop different strategies over time in responding to
the stimuli and that the transition from one strategy to the next is a
discrete event.  Before providing these illustrations, below we give a
formalization of dependent mixture models and a brief overview of the
DepmixS4 package that was developed to specify and fit such models.

@@

% outline
In this section we describe a class of models which are especially
suitable for describing and testing discrete change in (individual)
time series data.  The dependent mixture model is similar to, but
slightly different from, two other types of models that are in use for
modelling discrete change: the hidden and the latent Markov models.

% history of Markov models/LMM applications
Markov models have been used extensively in the social sciences; for
example, in analyzing language learning \cite{Miller1952,Miller1963},
in the analysis of paired associate learning \cite{Wickens1982}.  In
these models, the focus is on survey type data: a few repeated
measurements taken in a large sample.  \citeA{Langeheine1990} discuss
latent Markov models and their use in sociology and political science
(see also McCutcheon, 1987).  Latent transition models have been used
in studying development of math skills \citep{Collins1992} and in
medical applications \citep{Reboussin1998}; \citet{Kaplan2008}
provides an overview of such models, that are called stage-sequential
models in the developmental psychology literature.

@@

% hidden Markov models
Hidden Markov models (HMM) tend to be used in the analysis of long
univariate (individual) timeseries.  For example, HMMs are the model
of choice in speech recognition applications \cite{Rabiner1989}.  In
biology, HMMs are used to analyze DNA sequences \cite{Krogh1998} and
in econometric science, to analyze changes in stock market prices and
commodities \cite{Kim1994}.

% depmix model
The dependent mixture model that we propose here spans the range from
latent Markov models for few repeated measurements with many
participants to hidden Markov models for individual time series.  In
addition, the dependent mixture model includes multivariate responses.
The dependent mixture model consists of the following elements:
\begin{enumerate}
	\item $S$ is a collection of discrete states
	\item $S_{t} = \mat{A}S_{t-1}+\xi_{t}, \mat{A}$, a transition matrix
	\item $O_{t} = \mat{B}(S_{t}) + \zeta_{t}, \mat{B}$,  an observation density
\end{enumerate}
Here $\xi_{t}$ and $\zeta_{t}$ are independent error processes 
\cite{Elliott1995}.

% state space
The state space, which is a set of discrete states, captures the
different states that the learning or developmental process under
consideration can be in.  In the balance scale example mentioned
above, children are applying one of two possible strategies in
responding to the items.  The states are characterized by their
corresponding observation densities.  Using for example Rule 1 in the
balance scale task leads to correct answers on some items and
incorrect answers on others.  A different strategy may lead to correct
answers on some items and to guessing behavior on other items.  In
analyzing categorization learning data, in which participants learn to
categorize a set of objects, a typical initial state is that
participants are guessing because at the start of the task they have
no knowledge of which features are important in categorization.
Hence, the states in the state space represent knowledge states of the
participants, such that different knowledge states lead to different
observed behavior or responses.

@@

% log-likelihood
The log-likelihood of DMMs is usually computed by the
so-called forward-backward algorithm \citep{Baum1966,Rabiner1989}, or
rather by the forward part of this algorithm.  \citet{Lystig2002}
changed the forward algorithm in such a way as to allow computing the
gradients of the log-likelihood at the same time.  They start by
rewriting the likelihood as follows (for ease of exposition the
dependence on the model parameters is dropped here):
\begin{equation}
	L_{T} = Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) = \prod_{t=1}^{T}
Pr(\vc{O}_{t}|\vc{O}_{1},
	\ldots, \vc{O}_{t-1}),
	\label{condLike}
\end{equation}
where $Pr(\vc{O}_{1}|\vc{O}_{0}):=Pr(\vc{O}_{1})$. Note that for a
simple, i.e.\ observed, Markov chain these probabilities reduce to
$Pr(\vc{O}_{t}|\vc{O}_{1},\ldots,
\vc{O}_{t-1})=Pr(\vc{O}_{t}|\vc{O}_{t-1})$.
The log-likelihood can now be expressed as:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log[Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots,
\vc{O}_{t-1})].
	\label{eq:condLogl}
\end{equation}

To compute the log-likelihood, \citet{Lystig2002} define the following
(forward) recursion:
\begin{align}
	\phi_{1}(j) &:= Pr(\vc{O}_{1}, S_{1}=j) = \pi_{j} b_{j}(\vc{O}_{1})
	\label{eq:fwd1} \\
\begin{split}
	\phi_{t}(j) &:= Pr(\vc{O}_{t}, S_{t}=j|\vc{O}_{1}, \ldots,
\vc{O}_{t-1}) \\
	&= \sum_{i=1}^{N} [\phi_{t-1}(i)a_{ij}b_{j}(\vc{O}_{t})] \times
(\Phi_{t-1})^{-1},
	\label{eq:fwdt}
\end{split}
\end{align}
where $\Phi_{t}=\sum_{i=1}^{N} \phi_{t}(i)$. Combining
$\Phi_{t}=Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, \vc{O}_{t-1})$, and
equation~(\ref{eq:condLogl}) gives the following expression for the
log-likelihood:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log \Phi_{t}.
	\label{eq:logl}
\end{equation}

@@

% outline/abstract: this has redundant information, mostly discussed
% in detail below
\pkg{depmixS4} implements a general framework for defining and fitting
dependent mixture models in the R programming language \citep{R2008}.
This includes standard Markov models, latent/hidden Markov models, and
latent class and finite mixture distribution models.  The models can
be fitted on mixed multivariate data with multinomial and/or gaussian
distributions.

% design goals
The \pkg{depmixS4} package was motivated by the fact that Markovian
models are used commonly in the social sciences, but no comprehensive
package was available for fitting such models in R. Common programs
for Markovian models include Panmark \citep{Pol1996}, and for latent
class models Latent Gold \citep{Vermunt2003}.  Those programs are
lacking a number of features that were needed in our own research.  In
particular, \pkg{depmixS4} was designed to meet the following goals:
\begin{enumerate}

	\item to be able to fit transition models with covariates, i.e.,
	to have time-dependent transition matrices.

	\item to be able to include covariates in the prior or initial
	state probabilities of models.

	\item to allow for easy extensibility; in particular, to be able
	to add new response distributions, both univariate and
	multivariate, and similarly to be able to allow for the addition
	of other transition models, e.g., models for observations in 
	continuous time.

@@

Although \pkg{depmixS4} is designed to deal with longitudinal or time
series data, for say $T>100$, it can also handle the limit case with
$T=1$ in analyzing cross-sectional data.  In those cases, there are no
time dependencies between observed data, and the model reduces to a
finite mixture model \cite{McLachlan2000}, or a latent class model
\cite{McCutcheon1987}.  Although there are other specialized (R-)
packages to deal with mixture data, one specific feature that we
needed which is not available in other packages is the possibility to
include covariates on the prior probabilities of class membership.

The package is built using S4 classes (object oriented classes in R)
to allow easy extensibility (Chambers, 1998). A depmix model consists 
of the following: 
\begin{enumerate} 
	\item A model for the initial probabilities or prior 
	\item A list of transition models; one model for each row of the 
	transition matrix
	\item A list of response models for each of the states of the model; 
	this is a list of lists in the case of multivariate responses. 
\end{enumerate} 
Each of these are adressed briefly below.

\subsection{Transition and prior probabilities models}

% transition and initial state probs
Each row of the transition matrix and the vector of initial state 
probabilities is modelled as a baseline-category multinomial logistic 
model \cite<see>[chapter 7]{Agresti2002}. This means that covariates 
can be used as predictors in these models. In particular, the model 
for each multinomial is:
\begin{equation} 
	\pi_{i} = \frac{\exp(\alpha_{i}+\mathbf{\beta}^{`}_{i}\vc{x})}
				{1+\sum_{j=1}^{J-1}\exp(\alpha_{j}+\mathbf{\beta}^{`}_{j}\vc{x})},
\end{equation}
where $\pi_{i}$ is the probability (e.g.\ the probability of the
$i$-th initial state); the $\alpha_{i}$ are the category intercepts;
the $\mathbf{\beta}_{i}$ are the category regression coefficients;
$\vc{x}$ is a vector of covariates or predictors; $J$ is the number
of states; in this example, $J$ serves as the baseline-category,
meaning that $\alpha_{J}$ and $\mathbf{\beta}_{J}$ are zero.

\subsection{Response models}

% response models
Response models in \pkg{depmixS4} interface with the \code{glm} 
functions available in R and with the \pkg{nnet}-package for the 
multinomial logistic models. Normally distributed data can hence be 
modelled with direct effects included as well. For example, normal 
data are modelled as:
\begin{equation}
	O_{t}|S_{t} = \mu + \mathbf{\beta}^{`}\vc{x} + \epsilon_{t},
\end{equation}
where $\mu$ the (state-dependent) mean; $\vc{x}$ is a vector of
covariates or predictors; $\mathbf{\beta}$ is the vector of
(state-dependent) regression coefficients; $\epsilon_{t}$ is a
normally distributed error with (state-dependent) standard deviation
$\sigma$.

Multinomial responses are modeled by the same multinomial 
baseline-category logistic model as is used for the transition 
probabilities and the initial state probabilities. Other response 
functions from the generalized linear model family that can be used 
include: binomial logistic regression, poisson regression, gamma and 
inverse gaussian. There is also support for the multivariate normal 
distribution.

% MAARTEN: de multinomial logistic models maken weer gebruik van andere
% packages toch, kun je daar de referentie van geven?

@@

Parameters are estimated in \pkg{depmixS4} using the EM algorithm or
through the use of a general Newton-Raphson optimizer.  EM is used by
default in unconstrained models, but otherwise, direct optimization is
done using \pkg{Rdonlp2} \cite{Tamura2007,Spellucci2002}, because it
handles general linear (in)equality constraints, and optionally also
non-linear constraints.

@@

General issues
\begin{itemize}
	\item It is feasible to fit hidden Markov models in moderate length individual time series
	\item illustrations show results that are consistent with usual group based analyses AND 
	provide extra information on switching between strategies over the course of learning
	\item Many applications in experimental psychology: examples????
\end{itemize}

Specific insights from the illustrations: suggestions welcome

Possible future extensions in depmixS4
\begin{enumerate}
	\item richer measurement models, eg factor models, AR models etc
	\item richer transition models, eg continuous time measurement occasions
	\item explicit state durations
	\item identifiability of models
	\item model selection
\end{enumerate}
